---
title: "Run and build TRIALMON"
subtitle: "The complete R Shiny safety-screening console, every line"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    highlight-style: github-dark
    code-copy: true
    code-fold: false
execute:
  eval: false
  echo: true
---

## Run it

TRIALMON is a real R Shiny app. From the package root:

```r
install.packages(c("shiny", "bslib", "ggplot2"))   # once
shiny::runApp("sasr_monitoring_wiki/trialmon_shiny")
```

Click a point on the eDISH plot (or pick a participant in the sidebar) to open that
participant's evidence packet: the eDISH position, the QTcF tier, the adverse events,
the disposition, and the analyte trajectory. Every panel is a SCREENING lens. It flags
a position for the medical monitor / SRC / DSMB, it never reports a clinical number and
never adjudicates a DLT or a Hy's-Law case. Reported safety numbers come from the
validated safety database.

The whole tool is a few short files, shown here in full, nothing elided:

- **trialmon_shiny/R/logic.R** &mdash; the pure screening rules (no Shiny). The [visual guide](trialmon-guide.html), the [dashboard](trialmon-dashboard.html), and the unit tests all call these same functions.
- **trialmon_shiny/R/gen_data.R** &mdash; writes the synthetic CSVs so you can see the shape; swap them for a read-only safety-database export.
- **trialmon_shiny/app.R** &mdash; the UI and the reactive server.

The one idea to hold onto: the **rules** live in logic.R, the **screen** lives in app.R.
You can change what a screen means without touching the app, and change how it looks
without touching the rules.

## The data it reads

Three CSVs are the whole dataset. Open them directly, or edit them to point the app at
your study (keep the column names): [`participants.csv`](../trialmon_shiny/data/participants.csv)
(one row per participant, values as xULN multiples and ms, not measurements),
[`aes.csv`](../trialmon_shiny/data/aes.csv) (the adverse-event rows), and
[`cohorts.csv`](../trialmon_shiny/data/cohorts.csv) (the dose ladder). On a real study
you replace them with a read-only export from the validated safety database / ADaM
(ADLB, ADEG, ADAE), same columns. Nothing writes back.

## The complete screening core: R/logic.R

Every line of the screening logic. `trialmon_load()` reads the three CSVs;
`score_participants()` attaches the eDISH AND-gate and the QTcF tier to each participant;
`dlt_state()` rolls DLTs up per cohort for the 3+3 ladder; `trialmon_worklist()` builds
the ranked flag list; `trialmon_status()` and `trialmon_kpis()` summarise the whole
portfolio. Nothing here depends on Shiny, so you can test it in a plain console.

```{r logic, code=readLines("../trialmon_shiny/R/logic.R"), eval=FALSE, echo=TRUE}
```

Read the eDISH gate: `p$hys <- (p$alt > 3 | p$ast > 3) & p$tbl > 2`. That single line is
why a high ALT alone does not fire. Bilirubin must be up too. In the synthetic data
`CP101-009` (ALT 4.2x, TBili 1.6x) stays green because the second half of the `&` is
false, while `CP101-014` (ALT 5.6x, TBili 3.1x) lands in the RED quadrant. Change that
line and you change what "a Hy's-Law position" means, everywhere the guide, the
dashboard, the app, and the tests, all at once.

## The complete data generator: R/gen_data.R

Every line of the synthetic-data builder. It writes the three CSVs deterministically
from one seed, plants the signature illustrative positions (the RED `CP101-014`, the
near-miss `CP101-009`, the coordinated sub-threshold drift on `CP101-011`), and stops.
On a real study you delete this step and drop in the safety-database export instead.

```{r gendata, code=readLines("../trialmon_shiny/R/gen_data.R"), eval=FALSE, echo=TRUE}
```

## The complete app: app.R

Every line of the UI and server. Read it top to bottom; the walk-through below points
at the parts that matter.

```{r app, code=readLines("../trialmon_shiny/app.R"), eval=FALSE, echo=TRUE}
```

## How the reactivity works, part by part

Follow the data, not the layout. Here is the chain that makes a click on the scatter
open a participant's evidence packet.

**The whole portfolio is scored once, up front.** `tm <- trialmon_load("data")` reads the
CSVs and `P <- score_participants(tm$participants)` attaches the eDISH and QTcF columns.
The value boxes, the eDISH plot, the worklist, and the 3+3 ladder are all views of that
one scored frame, so they always agree.

**The signature scatter is a plot output with a click.** `output$edish` renders the log-log
ALT-vs-TBili scatter with the Hy's-Law quadrant shaded; the `plotOutput("edish", click="edish_click")`
in the UI turns every pixel of it into a click target.

**Clicking maps back to a participant.** The observer on `input$edish_click` calls
`nearPoints(P, input$edish_click, xvar="alt", yvar="tbl", maxpoints=1)` to find the
nearest point on the log-log plot, then `updateSelectInput(session, "pid", selected = np$id[1])`
sets the selected participant. You never wire the click to the packet by hand; changing
`input$pid` is enough.

**The selected participant drives the packet.** `sel <- reactive(P[P$id == input$pid, ])`
is the one reactive everything on the detail side depends on. When `input$pid` changes,
`sel()` recomputes, and `output$packet` (the sidebar evidence packet) and `output$traj`
(the analyte trajectory) both re-render for the new participant.

**The screens read the scored frame, not each other.** `output$worklist` calls
`trialmon_worklist(tm)`, `output$ladder` calls `dlt_state(tm)`, and the value boxes read
`trialmon_kpis(tm)`. Each is a pure function of the data, so there is no hidden order of
operations to get wrong.

## Verify it without a browser

The screening rules are covered by a zero-dependency test you run with `Rscript`. It is
the same file the guide and dashboard trust, and it pins the AND-gate directly: a high
ALT with normal bilirubin must NOT be flagged as a Hy's-Law position.

```{r tests, code=readLines("../trialmon_tests/test-logic.R"), eval=FALSE, echo=TRUE}
```

Run it:

```r
Rscript sasr_monitoring_wiki/trialmon_tests/test-logic.R   # prints: TRIALMON logic: all tests passed
```

## Make it yours

- **Change a threshold:** edit `qtc_sev()` or the eDISH gate in `R/logic.R`. Everything that shows a screen updates, because they all call that one function.
- **Change the data:** swap the three CSVs for a read-only safety-database / ADaM export with the same columns.
- **Add a screen:** add a rule to `R/logic.R` and a `trialmon_worklist()` entry, then a matching `render*` in `app.R`.
- **Add a panel:** another `card()` in the UI and a matching `renderPlot`/`renderUI` in the server.

The same skeleton, a logic core plus scored frame plus pure summary functions plus
outputs, builds the [lifecycle monitor](../../lifecycle_dashboard/quarto/lifecycle-build.html)
and the [find-hours app](../../budget_reconciliation/quarto/hours-build.html) too.
