---
title: "IRT / RTSM UAT Snapshot"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    embed-resources: true
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(ggplot2); library(knitr)
source("../shiny/R/logic.R")
d <- uat_load("../templates")
g <- uat_gate(d$tests, d$defects, d$trace)
s <- uat_summary(d$tests, d$defects, d$trace)
```

::: {.callout-note appearance="simple"}
**Read-only shareable snapshot**, computed from the validated UAT data ([`test_scripts.csv`](../templates/test_scripts.csv), [`defect_log.csv`](../templates/defect_log.csv), [`traceability_matrix.csv`](../templates/traceability_matrix.csv)). It does not connect to the IRT. The interactive tool is the [Run &amp; build the Shiny app](uat-build.html) page, and the [visual guide](uat-guide.html) explains it. Operations-only, synthetic, no PHI. We say *participant*, not *subject*.
:::

```{r kpis, results='asis'}
box <- function(label, value, col) sprintf(
  '<div style="flex:1 1 120px;background:#222630;border:1px solid #3a4157;border-radius:10px;padding:14px 16px"><div style="font-size:30px;font-weight:700;color:%s;font-family:monospace;line-height:1">%s</div><div style="font-size:11px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em;margin-top:6px">%s</div></div>',
  col, value, label)
cat('<div style="display:flex;flex-wrap:wrap;gap:10px;margin:6px 0 18px">')
cat(box("Test scripts", s$total, "#e9edf7"))
cat(box("Passed", s$pass, "#43c47d"))
cat(box("Failed", s$fail, "#e8657a"))
cat(box("Open defects", s$open_def, "#e0a64c"))
cat(box("Requirements traced", paste0(s$cov, "%"), "#3fd0c9"))
cat('</div>')
```

:::: {.grid}

::: {.g-col-12 .g-col-md-5}
#### Go-live sign-off gate

```{r gate, results='asis'}
tick <- function(ok, txt) sprintf(
  '<div style="margin:6px 0;font-size:15px;color:#e9edf7"><span style="color:%s;font-weight:700;margin-right:6px">%s</span>%s</div>',
  if (ok) "#43c47d" else "#e8657a", if (ok) "&#10004;" else "&#10008;", txt)
bg <- if (g$go) "#152b1e" else "#2b1519"; bd <- if (g$go) "#43c47d" else "#e8657a"
cat(sprintf('<div style="background:%s;border:1px solid %s;border-radius:12px;padding:16px 18px">', bg, bd))
cat(sprintf('<div style="font-size:22px;font-weight:700;margin-bottom:10px;color:#e9edf7">%s</div>', g$verdict))
cat(tick(g$crit_fail == 0, sprintf("<b>%d</b> critical test(s) not yet passing", g$crit_fail)))
cat(tick(g$open_cm == 0,  sprintf("<b>%d</b> open critical/major defect(s)", g$open_cm)))
cat(tick(g$trace_gap == 0, sprintf("<b>%d</b> requirement(s) without a test", g$trace_gap)))
cat('<div style="font-size:12px;color:#94a1c0;margin-top:12px">Conjunctive: all three must hold. The gate packages the evidence, a named human signs.</div></div>')
```
:::

::: {.g-col-12 .g-col-md-7}
#### Execution by area

```{r plot, fig.width=7, fig.height=4.2}
df <- uat_area_counts(d$tests)
ggplot(df, aes(Freq, Area, fill = Status)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = c("Pass"="#43c47d","Fail"="#e8657a","Blocked"="#e0a64c",
    "In Progress"="#5b8cff","Not Run"="#6b7794"), drop = FALSE) +
  labs(x = "Tests", y = NULL, fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", plot.background = element_rect(fill="#222630", colour=NA),
        panel.background = element_rect(fill="#222630", colour=NA),
        text = element_text(colour="#e9edf7"), axis.text = element_text(colour="#c5cee4"),
        panel.grid = element_line(colour="#3a4157"),
        legend.key = element_rect(fill="#222630", colour=NA))
```
:::

::::

#### Test scripts

```{r table}
tt <- d$tests[, c("TestID","Area","Priority","TestTitle","Status")]
kable(tt, col.names = c("Test","Area","Priority","Title","Status"), row.names = FALSE)
```
