---
title: "IRT / RTSM User Acceptance Testing"
subtitle: "Confirming the build before the first participant is randomized"
format:
  revealjs:
    theme: dark
    embed-resources: true
    slide-number: true
    transition: fade
    fig-width: 9
    fig-height: 4.2
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(ggplot2); library(knitr)
source("../shiny/R/logic.R")
d <- uat_load("../templates")
g <- uat_gate(d$tests, d$defects, d$trace)
s <- uat_summary(d$tests, d$defects, d$trace)
```

## What a UAT is

You confirm, in a test environment, that the vendor's randomization and drug-supply build does **exactly** what your specs say, before a single real participant is randomized.

```{mermaid}
flowchart LR
  URS[URS + specs] --> TC[Test cases]
  TC --> EX[Execute in<br/>staging IRT]
  EX --> DEF[Defects]
  DEF --> EX
  EX --> TR[Traceability]
  TR --> GATE{{Go-live gate}}
  GATE --> PROD[Production]
```

Operations-only, synthetic, no PHI. We say *participant*, not *subject*.

## Where it runs: the systems

- **Staging IRT / RTSM** (vendor): the web app you operate in a browser, plus the approved test list.
- **You, the biostatistician:** execute each case, capture dated screenshot evidence, and transcribe results into your workbook.
- **Vendor defect tracker:** log failures; the vendor promotes a fixed build back to staging for retest.
- **eTMF:** the signed sign-off package is filed as the record.

The workbook does not connect to the IRT. You transcribe on purpose, so the record is attributable and stands alone for an auditor.

## The live gate, computed now

```{r results='asis'}
tick <- function(ok, txt) sprintf('<div style="margin:6px 0;font-size:26px"><span style="color:%s;font-weight:700">%s</span> %s</div>',
  if (ok) "#43c47d" else "#e8657a", if (ok) "&#10004;" else "&#10008;", txt)
bg <- if (g$go) "#1e3a28" else "#3a1e24"; bd <- if (g$go) "#43c47d" else "#e8657a"
cat(sprintf('<div style="background:%s;border:1px solid %s;border-radius:12px;padding:20px 24px;max-width:820px">', bg, bd))
cat(sprintf('<div style="font-size:34px;font-weight:700;margin-bottom:10px;color:#e9edf7">%s</div>', g$verdict))
cat(tick(g$crit_fail == 0, sprintf("<b>%d</b> critical test(s) not yet passing", g$crit_fail)))
cat(tick(g$open_cm == 0,  sprintf("<b>%d</b> open critical/major defect(s)", g$open_cm)))
cat(tick(g$trace_gap == 0, sprintf("<b>%d</b> requirement(s) without a test", g$trace_gap)))
cat('</div>')
```

The gate is **conjunctive**: all critical pass **AND** 0 open critical/major defects **AND** 100% traceability.

## Snapshot

```{r results='asis'}
box <- function(label, value, col) sprintf(
  '<div style="flex:1;min-width:150px;background:#222630;border:1px solid #3a4157;border-radius:12px;padding:18px 20px;margin:6px"><div style="font-size:40px;font-weight:700;color:%s;font-family:monospace">%s</div><div style="font-size:15px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em">%s</div></div>', col, value, label)
cat('<div style="display:flex;flex-wrap:wrap;gap:4px">')
cat(box("Test scripts", s$total, "#e9edf7"))
cat(box("Passed", s$pass, "#43c47d"))
cat(box("Failed", s$fail, "#e8657a"))
cat(box("Open defects", s$open_def, "#e0a64c"))
cat(box("Traced", paste0(s$cov, "%"), "#3fd0c9"))
cat('</div>')
```

Every number here comes from the validated data, never from an assistant.

## Execution by area

```{r}
df <- uat_area_counts(d$tests)
ggplot(df, aes(Freq, Area, fill = Status)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = c("Pass"="#43c47d","Fail"="#e8657a","Blocked"="#e0a64c",
    "In Progress"="#5b8cff","Not Run"="#6b7794"), drop = FALSE) +
  labs(x = "Tests", y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top", plot.background = element_rect(fill="#222630", colour=NA),
        panel.background = element_rect(fill="#222630", colour=NA),
        text = element_text(colour="#e9edf7"), axis.text = element_text(colour="#c5cee4"),
        panel.grid = element_line(colour="#3a4157"),
        legend.key = element_rect(fill="#222630", colour=NA))
```

## How to write an executable test

**RAND-02, stratified allocation.** A good test has a specific, verifiable expected result, not "works correctly".

- **Precondition:** test list loaded; the <70 kg stratum, block 1, has an unused slot.
- **Steps:** register participant 101-001 at 68.0 kg, randomize, record kit + confirmation, open the allocation report as the unblinded statistician.
- **Expected:** allocated to the **next unused <70 kg / block-1 entry**; arm, sequence, and kit match the list; the investigator sees a **kit ID only**, not the arm.
- **Verify:** compare cell-for-cell across every stratum and both arms, not one happy-path participant.

## Severity drives the gate

```{r results='asis'}
cat('
| Severity | Meaning | Blocks go-live? |
|---|---|---|
| **Critical** | Could break the blind, mis-randomize, mis-dose, or corrupt integrity | **Yes** |
| **Major** | Significant functional error in a key workflow (resupply, wrong kit, missing alert) | **Yes, until fixed** |
| **Minor** | Cosmetic / non-functional (label typo, missing site number) | No; fix or waive with rationale |
')
```

One open Critical or Major defect holds the gate. That is the whole point: severity, not count, decides.

## Run it, and build it

The interactive tool is a real R Shiny app you own:

```r
shiny::runApp("uat_irt/shiny")
# needs: install.packages(c("shiny","bslib","DT","ggplot2"))
```

Mark a test Pass/Fail, close a defect, and the KPIs + go-live gate recompute live. The gate math lives in `shiny/R/logic.R` (~50 lines), shared by the app, these slides, and the unit tests. Change the CSV to your study's tests and it is your tracker.

## Thank you

- **Explore / run:** the [Shiny app](../shiny)
- **Share (no R):** the [dashboard](uat-dashboard.html)
- **Skim:** the [one-page cheat-sheet](uat-onepager.pdf)
- **Read:** the [full guide](uat-guide.html)

*Synthetic data throughout. UAT sign-off is a named-human decision.*
