---
title: "Run and build the UAT tracker"
subtitle: "The complete R Shiny app, every line"
format:
  html:
    theme: darkly
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    highlight-style: github-dark
    code-copy: true
    code-fold: false
execute:
  eval: false
  echo: true
---

## Run it

The interactive tool is a real R Shiny app. From the package root:

```r
install.packages(c("shiny", "bslib", "DT", "ggplot2"))   # once
shiny::runApp("uat_irt/shiny")
```

Select a test and mark it Pass/Fail, close a defect, and the KPIs and the go-live
sign-off gate recompute live. Download your results to CSV from the sidebar to drop
into the [Excel workbook](../UAT_IRT_RTSM_Workbook.xlsx) and the eTMF.

The whole app is two files, shown here in full, nothing elided:

- **shiny/R/logic.R** &mdash; the pure gate and KPI logic (no Shiny). The [visual guide](uat-guide.html), the [dashboard](uat-dashboard.html), and the unit tests call these same functions.
- **shiny/app.R** &mdash; the UI and the reactive server.

The one idea to hold onto: the **rules** live in `logic.R`, the **screen** lives in `app.R`. You can change what the gate means without touching the app, and change how it looks without touching the rules.

## The data it reads

Three CSVs are the whole dataset. Edit them to point the app at your study (keep the column names): [`test_scripts.csv`](../templates/test_scripts.csv), [`defect_log.csv`](../templates/defect_log.csv), [`traceability_matrix.csv`](../templates/traceability_matrix.csv).

## The complete logic core: shiny/R/logic.R

Every line of the logic. `uat_load()` reads the three CSVs; `uat_gate()` is the
conjunctive go-live rule; `uat_summary()` is the KPI row; `uat_area_counts()` shapes
the by-area bar. Nothing here depends on Shiny, so you can test it in a plain console.

```{r logic, code=readLines("../shiny/R/logic.R"), eval=FALSE, echo=TRUE}
```

Read the gate: it counts three things and requires all three to be zero. A Critical
test not yet Pass, an open Critical or Major defect, or a requirement with no test,
any one of them holds the gate. That is the whole meaning of "conjunctive", written
as `go <- crit_fail == 0 && open_cm == 0 && trace_gap == 0`. Change that line and you
have changed what "ready" means, everywhere, at once.

## The complete app: shiny/app.R

Every line of the UI and server. Read it top to bottom; the walk-through below points
at the parts that matter.

```{r app, code=readLines("../shiny/app.R"), eval=FALSE, echo=TRUE}
```

## How the reactivity works, part by part

Follow the data, not the layout. Here is the chain that makes a click flip the gate.

**The editable state lives in one place.** `rv <- reactiveValues(tests = seed$tests, defects = seed$defects, trace = seed$trace)`
holds everything the user can change. Every output reads from `rv`, so when `rv` changes,
everything that depends on it recomputes. That is the entire trick.

**The rules are reactive expressions.** `gate <- reactive(uat_gate(rv$tests, rv$defects, rv$trace))`
and `summ <- reactive(uat_summary(rv$tests, rv$defects, rv$trace))` re-run automatically whenever any of the three
frames in `rv` changes. You never call the gate by hand.

**Selecting a row shows its detail.** The tests table is a `DT` with single-row selection.
`input$tests_rows_selected` gives the selected row; `output$detail` renders that test's
steps, expected result, and a `radioButtons` control set to its current status.

**Marking a test mutates the state.** The observer on `input$set_status` writes the new
status into `rv$tests` for the selected row. That write is what makes the gate flip, because
`gate()` depends on `rv$tests`. You do not recompute anything yourself; changing `rv`
triggers the reactive that does.

**Closing a defect does the same.** The defect table is editable on its Status column; the
observer on `input$defects_cell_edit` accepts only "Open" or "Closed", writes it into
`rv$defects`, and the gate recomputes.

**Everything else just reads the reactives.** The value boxes read `summ()`, the gate card
reads `gate()`, the by-area plot reads `rv$tests`, and the download handler writes `rv$tests`
to CSV so results round-trip to Excel and the eTMF.

## Verify it without a browser

You can drive the reactive graph headless, which is how this app was checked before it
was ever opened in a browser:

```r
shiny::testServer("uat_irt/shiny", {
  # mark the failing critical test as Pass and watch the count drop
  session$setInputs(
    tests_rows_selected = which(rv$tests$TestID == "RAND-04"),
    set_status = "Pass"
  )
  gate()$crit_fail            # one lower than before

  # close the two open critical/major defects
  session$setInputs(defects_cell_edit = list(row = which(rv$defects$DefectID == "DEF-01"),
                                             col = 2, value = "Closed"))
  gate()$open_cm              # drops toward zero
})
```

## The unit tests: shiny is not involved

The gate logic is covered by a zero-dependency test you run with `Rscript`. It is the
same file the guide and dashboard trust:

```{r tests, code=readLines("../tests/test-logic.R"), eval=FALSE, echo=TRUE}
```

Run it:

```r
Rscript uat_irt/tests/test-logic.R      # prints: UAT logic: all tests passed
```

## Make it yours

- Change the rule: edit `uat_gate()` in `shiny/R/logic.R`, for example to allow a Minor-defect waiver. Nothing else changes, because everything calls that one function.
- Change the data: edit the three CSVs in `templates/`.
- Add a column: add it to the CSV and to the `datatable()` column list in `app.R`.
- Add a panel: another `card()` in the UI and a matching `render*` in the server.

The same skeleton, a logic core plus `reactiveValues` plus reactive rules plus outputs,
builds the [find-hours app](../../budget_reconciliation/quarto/hours-build.html) and the
four dashboards too.
