---
title: "Study Lifecycle Monitor - Snapshot"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    embed-resources: true
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(ggplot2); library(knitr)
source("../shiny/R/logic.R")
d  <- lifecycle_load("../shiny/data")
st <- score_portfolio(d)
k  <- lifecycle_kpis(st, d)
```

::: {.callout-note appearance="simple"}
**Read-only shareable snapshot**, computed from the portfolio ([`studies.csv`](../shiny/data/studies.csv), [`milestones.csv`](../shiny/data/milestones.csv), [`deliverables.csv`](../shiny/data/deliverables.csv), [`query_trend.csv`](../shiny/data/query_trend.csv)). The interactive tool is the [run-and-build page](lifecycle-build.html), and the [visual guide](lifecycle-guide.html) explains the scoring. Synthetic, no PHI, no participant-level data.
:::

```{r kpis, results='asis'}
box <- function(label, value, col) sprintf(
  '<div style="flex:1 1 120px;background:#222630;border:1px solid #3a4157;border-radius:10px;padding:14px 16px"><div style="font-size:28px;font-weight:700;color:%s;font-family:monospace;line-height:1">%s</div><div style="font-size:11px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em;margin-top:6px">%s</div></div>',
  col, value, label)
cat('<div style="display:flex;flex-wrap:wrap;gap:10px;margin:6px 0 18px">')
cat(box("Active studies", k$n, "#e9edf7"))
cat(box("At risk", k$at_risk, "#e8657a"))
cat(box("On watch", k$watch, "#e0a64c"))
cat(box("Open queries", k$open_q, "#e9edf7"))
cat(box("Avg deliverables", paste0(k$avg_del,"%"), "#3fd0c9"))
cat(box("Next DB lock", ifelse(is.na(k$next_dbl_days),"-",paste0(k$next_dbl_days,"d")), "#5b8cff"))
cat('</div>')
```

:::: {.grid}

::: {.g-col-12 .g-col-md-7}
#### Portfolio risk heatmap

```{r heat, fig.width=7, fig.height=4.2}
dims <- c("Timeline","Data","TLF","Resource")
hl <- do.call(rbind, lapply(dims, function(dm) data.frame(id=st$id, dim=dm, value=st[[dm]])))
hl$id <- factor(hl$id, levels=rev(st$id)); hl$dim <- factor(hl$dim, levels=dims)
ggplot(hl, aes(dim, id, fill=value)) +
  geom_tile(colour="#222630", linewidth=1.5) +
  geom_text(aes(label=value), colour="#0b1020", fontface="bold", size=4) +
  scale_fill_gradientn(colours=c("#43c47d","#e0a64c","#e8657a"), limits=c(0,100)) +
  labs(x=NULL, y=NULL) + theme_minimal(base_size=12) +
  theme(legend.position="none", panel.grid=element_blank(),
        plot.background=element_rect(fill="#222630", colour=NA),
        panel.background=element_rect(fill="#222630", colour=NA),
        text=element_text(colour="#e9edf7"), axis.text=element_text(colour="#c5cee4"))
```
:::

::: {.g-col-12 .g-col-md-5}
#### Early-warning signals (ranked)

```{r signals, results='asis'}
sg <- lifecycle_signals(st, d)
for (j in seq_len(min(8, nrow(sg)))) {
  r <- sg[j, ]; col <- if (r$sev == "red") "#e8657a" else "#e0a64c"
  cat(sprintf('<div style="display:flex;gap:8px;padding:6px 0;border-bottom:1px solid #2b3350"><div style="flex:0 0 8px;height:8px;border-radius:50%%;background:%s;margin-top:6px"></div><div><div style="font-size:13px;color:#e9edf7"><b>%s</b> - %s</div><div style="font-size:12px;color:#94a1c0">%s</div></div></div>', col, r$who, r$title, r$msg))
}
```
:::

::::

#### Portfolio open-query trend (8 weeks)

```{r trend, fig.width=10, fig.height=2.6}
agg <- aggregate(open_queries ~ week, d$query_trend, sum)
ggplot(agg, aes(week, open_queries)) +
  geom_line(colour="#5b8cff", linewidth=1.2) + geom_point(colour="#5b8cff", size=2) +
  labs(x="Week", y="Open queries (portfolio)") + theme_minimal(base_size=12) +
  theme(plot.background=element_rect(fill="#222630", colour=NA),
        panel.background=element_rect(fill="#222630", colour=NA),
        text=element_text(colour="#e9edf7"), axis.text=element_text(colour="#c5cee4"),
        panel.grid=element_line(colour="#3a4157"))
```

#### Studies, ranked by overall risk

```{r table}
kable(st[, c("id","phase","overall","Timeline","Data","TLF","Resource","rag")],
      col.names=c("Study","Phase","Overall","Timeline","Data","TLF","Resource","RAG"), row.names=FALSE)
```
