---
title: "Run and build the Study Lifecycle Monitor"
subtitle: "The complete R Shiny console, every line"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    highlight-style: github-dark
    code-copy: true
    code-fold: false
execute:
  eval: false
  echo: true
---

## Run it

The interactive tool is a real R Shiny app. From the package root:

```r
install.packages(c("shiny", "bslib", "DT", "ggplot2"))   # once
shiny::runApp("lifecycle_dashboard/shiny")
```

Pick a study, then move the what-if sliders (open queries, double-programmed %) and
watch that study's row in the risk heatmap re-colour. The scoring is transparent, so
every number is defensible.

The whole console is three R files, shown here in full, nothing elided:

- **shiny/R/logic.R** &mdash; the pure scoring and signal logic (no Shiny). The [visual guide](lifecycle-guide.html), the [dashboard](lifecycle-dashboard.html), and the unit tests call these same functions.
- **shiny/R/gen_data.R** &mdash; the deterministic synthetic portfolio, the four CSVs the app reads.
- **shiny/app.R** &mdash; the UI and the reactive server.

The one idea to hold onto: the **rules** live in logic.R, the **screen** lives in app.R.
You can change what "risk" means without touching the app, and change how it looks
without touching the rules.

## The data it reads

Four CSVs are the whole dataset. Edit them to point the app at your study (keep the
column names): [`studies.csv`](../shiny/data/studies.csv), [`milestones.csv`](../shiny/data/milestones.csv), [`deliverables.csv`](../shiny/data/deliverables.csv), [`query_trend.csv`](../shiny/data/query_trend.csv). On a real study you replace them with a read-only CTMS / EDC export carrying the same columns; nothing writes back to a source system.

## The complete logic core: shiny/R/logic.R

Every line of the scoring. `lifecycle_load()` reads the four CSVs; `score_study()` is
the transparent per-study blend; `score_portfolio()` attaches the score columns and the
RAG band; `lifecycle_kpis()` is the KPI row; `lifecycle_signals()` ranks the
early-warning signals. Nothing here depends on Shiny, so you can test it in a plain
console.

```{r logic, code=readLines("../shiny/R/logic.R"), eval=FALSE, echo=TRUE}
```

Read the score: `score_study()` builds four sub-scores, then blends them as
`overall <- clamp(0.30 * timeline + 0.28 * data + 0.27 * tlf + 0.15 * resource)`, and
`score_portfolio()` turns that into a RAG band with
`st$rag <- ifelse(st$overall > 66, "red", ifelse(st$overall > 33, "amber", "green"))`.
Change those two lines and you have changed what "at risk" means, everywhere, at once.

## The complete data generator: shiny/R/gen_data.R

Every line of the synthetic portfolio. `gen_portfolio()` writes the four CSVs
deterministically (fixed seed), and forces `CP-108` into a clearly at-risk state so the
red gate and the signals are exercised. Swap this out for a real export and the app does
not change.

```{r gendata, code=readLines("../shiny/R/gen_data.R"), eval=FALSE, echo=TRUE}
```

## The complete app: shiny/app.R

Every line of the UI and server. Read it top to bottom; the walk-through below points
at the parts that matter.

```{r app, code=readLines("../shiny/app.R"), eval=FALSE, echo=TRUE}
```

## How the reactivity works, part by part

Follow the data, not the layout. Here is the chain that makes a slider re-colour the
heatmap.

**The editable state lives in one place.** `rv <- reactiveValues(studies = seed$studies)`
holds the one frame the user can change. Every output reads from a reactive built on
`rv`, so when `rv$studies` changes, everything that depends on it recomputes. That is the
entire trick.

**The what-if writes back to that state.** The observer on `list(input$openq, input$dblprog)`
finds the selected study's row with `which(rv$studies$id == input$study)` and writes the
slider values into `rv$studies$open_queries[i]` and `rv$studies$dbl_prog_pct[i]`. A
second observer on `input$study` keeps the sliders in sync when you switch studies, and
`input$reset` restores `seed$studies`.

**The rules are reactive expressions.** The `data_now` reactive reassembles the portfolio
(studies from `rv$studies`, plus the seed milestones, deliverables, and query trend),
`st <- reactive(score_portfolio(data_now()))` re-scores it, and
`kpi <- reactive(lifecycle_kpis(st(), data_now()))` rolls it up. These re-run automatically
whenever `rv$studies` changes. You never call the scorer by hand.

**Every output just reads the reactives.** The value boxes read `kpi()`, the heatmap reads
`st()` through `heat_long()`, the ranked signals read `lifecycle_signals(st(), data_now())`,
and the per-study milestone track and TLF funnel read `rv$studies` for the selected `input$study`.

**Clicking the heatmap selects a study.** The observer on `input$heat_click` rounds the click
y-position to a row index into `rev(x$id)` and calls `updateSelectInput()` so the detail
panels follow. Nothing recomputes by hand; changing `rv` (or the selection) triggers the
reactive that does.

## Verify it without a browser

You can drive the reactive graph headless, which is how this app was checked before it
was ever opened in a browser:

```r
shiny::testServer("lifecycle_dashboard/shiny", {
  # push a study's open queries to the top of the slider and watch its risk rise
  session$setInputs(study = "CP-108", openq = 150)
  st()$overall[st()$id == "CP-108"]     # rises as open queries climb
})
```

## The unit tests: shiny is not involved

The scoring is covered by a zero-dependency test you run with `Rscript`. It is the same
file the guide and dashboard trust:

```{r tests, code=readLines("../tests/test-logic.R"), eval=FALSE, echo=TRUE}
```

Run it:

```r
Rscript lifecycle_dashboard/tests/test-logic.R   # prints: Lifecycle logic: all tests passed
```

## Make it yours

- **Change the rule:** edit `score_study()` (or the signal rules in `lifecycle_signals()`) in `shiny/R/logic.R`. Nothing else changes, because everything calls those functions.
- **Change the data:** swap the four CSVs in `shiny/data/` for a CTMS export with the same columns.
- **Add a dimension:** add a sub-score in `score_study()`, add its column to `score_portfolio()`, and add it to the heatmap's `dims` vector.
- **Add a panel:** another `card()` in the UI and a matching `render*` in the server.

The same skeleton, a logic core plus `reactiveValues` plus reactive rules plus outputs,
builds the [UAT tracker](../../uat_irt/quarto/uat-build.html) and the
[find-hours app](../../budget_reconciliation/quarto/hours-build.html) too.
