---
title: "Find Hours - Reconciliation Snapshot"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    embed-resources: true
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(ggplot2); library(knitr)
source("../shiny/R/logic.R")
d  <- hours_load()
m  <- hours_metrics(d$ledger, d$contingency)
f45 <- find_hours(45, d$ledger, d$contingency)
```

::: {.callout-note appearance="simple"}
**Read-only shareable snapshot**, computed from the effort ledger [`project_hours_ledger.csv`](../templates/project_hours_ledger.csv). It connects to no real budget or timesheet. The interactive tool is the [Run &amp; build the Shiny app](hours-build.html) page, and the [visual guide](hours-guide.html) explains it. Operations-only, synthetic, hours / effort only (no dollars, rates, or headcount). We say *participant*, not *subject*.
:::

```{r kpis, results='asis'}
box <- function(label, value, col) sprintf(
  '<div style="flex:1 1 120px;background:#222630;border:1px solid #3a4157;border-radius:10px;padding:14px 16px"><div style="font-size:30px;font-weight:700;color:%s;font-family:monospace;line-height:1">%s</div><div style="font-size:11px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em;margin-top:6px">%s</div></div>',
  col, value, label)
cat('<div style="display:flex;flex-wrap:wrap;gap:10px;margin:6px 0 18px">')
cat(box("Planned (h)",      m$planned,     "#e9edf7"))
cat(box("Spent to date (h)", m$actual,     "#5b8cff"))
cat(box("Forecast EAC (h)",  m$eac,        "#e0a64c"))
cat(box("Contingency (h)",   m$contingency, "#3fd0c9"))
cat(box("Net findable (h)",  m$findable,   "#43c47d"))
cat('</div>')
```

:::: {.grid}

::: {.g-col-12 .g-col-md-7}
#### Planned vs forecast (EAC) by deliverable

```{r plot, fig.width=7, fig.height=4.4}
long <- rbind(
  data.frame(deliverable = d$ledger$deliverable, hours = d$ledger$planned, kind = "Planned"),
  data.frame(deliverable = d$ledger$deliverable, hours = d$ledger$eac,     kind = "Forecast EAC"))
long$deliverable <- factor(long$deliverable, levels = rev(d$ledger$deliverable))
ggplot(long, aes(hours, deliverable, fill = kind)) +
  geom_col(width = 0.7, position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = c("Planned"="#5b8cff","Forecast EAC"="#e0a64c")) +
  labs(x = "Hours", y = NULL, fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top", plot.background = element_rect(fill="#222630", colour=NA),
        panel.background = element_rect(fill="#222630", colour=NA),
        text = element_text(colour="#e9edf7"), axis.text = element_text(colour="#c5cee4"),
        panel.grid = element_line(colour="#3a4157"),
        legend.key = element_rect(fill="#222630", colour=NA))
```
:::

::: {.g-col-12 .g-col-md-5}
#### A 45 h ask, sourced

```{r verdict, results='asis'}
f <- f45
bg <- if (f$absorbable) "#152b1e" else "#2b1519"; bd <- if (f$absorbable) "#43c47d" else "#e8657a"
row <- function(lbl, val, col) sprintf('<div style="margin:5px 0;color:#e9edf7"><span style="font-family:monospace;color:%s;font-weight:700;margin-right:8px">%2d h</span>%s</div>', col, val, lbl)
cat(sprintf('<div style="background:%s;border:1px solid %s;border-radius:12px;padding:16px 18px">', bg, bd))
cat(sprintf('<div style="font-size:20px;font-weight:700;margin-bottom:10px;color:#e9edf7">Ask 45 h &rarr; %s</div>', f$verdict))
cat(row("from under-run slack",     f$from_slack, "#43c47d"))
cat(row("from contingency reserve", f$from_cont,  "#3fd0c9"))
cat(row("residual (change order)",  f$shortfall,  if (f$shortfall > 0) "#e8657a" else "#94a1c0"))
cat(sprintf('<div style="font-size:12px;color:#94a1c0;margin-top:10px">Net findable = (&Sigma; planned + contingency) &minus; &Sigma; EAC = <b>%d h</b>. Contingency left after this ask: %d h.</div></div>', m$findable, f$contingency_left))
```
:::

::::

#### Effort ledger (hours)

```{r ledger}
lg <- d$ledger; lg$slack <- pmax(0, lg$planned - lg$eac)
kable(lg[, c("deliverable","planned","actual","pct","eac","slack")],
      col.names = c("Deliverable","Planned","Actual","% done","EAC","Under-run slack"),
      row.names = FALSE, align = "lrrrrr")
```
