---
title: "Find Hours — reconciliation snapshot"
format:
  dashboard:
    theme: darkly
    embed-resources: true
    nav-buttons: []
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(ggplot2); library(knitr)
source("../shiny/R/logic.R")
d   <- hours_load()
m   <- hours_metrics(d$ledger, d$contingency)
f45 <- find_hours(45, d$ledger, d$contingency)
```

::: {.callout-note appearance="simple"}
Shareable snapshot, computed from the effort ledger (hours only, synthetic). The [Shiny app](../shiny) is the interactive version. Every number comes from the validated ledger, never from an assistant.
:::

## Row {height=22%}

```{r}
#| content: valuebox
#| title: "Planned (h)"
#| color: "#222630"
list(value = m$planned)
```

```{r}
#| content: valuebox
#| title: "Spent to date (h)"
#| color: "#5b8cff"
list(value = m$actual)
```

```{r}
#| content: valuebox
#| title: "Forecast EAC (h)"
#| color: "#e0a64c"
list(value = m$eac)
```

```{r}
#| content: valuebox
#| title: "Contingency (h)"
#| color: "#3fd0c9"
list(value = m$contingency)
```

```{r}
#| content: valuebox
#| title: "Net findable (h)"
#| color: "#43c47d"
list(value = m$findable)
```

## Row {height=48%}

### Column {width=58%}

```{r}
#| title: "Planned vs forecast EAC, by deliverable"
df <- d$ledger[, c("deliverable","planned","eac")]
long <- rbind(
  data.frame(deliverable = df$deliverable, hours = df$planned, kind = "Planned"),
  data.frame(deliverable = df$deliverable, hours = df$eac,     kind = "Forecast EAC"))
long$deliverable <- factor(long$deliverable, levels = rev(df$deliverable))
ggplot(long, aes(hours, deliverable, fill = kind)) +
  geom_col(width = 0.7, position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = c("Planned"="#5b8cff","Forecast EAC"="#e0a64c")) +
  labs(x = "Hours", y = NULL, fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.background  = element_rect(fill="#222630", colour=NA),
        panel.background = element_rect(fill="#222630", colour=NA),
        text = element_text(colour="#e9edf7"),
        axis.text = element_text(colour="#c5cee4"),
        panel.grid = element_line(colour="#3a4157"))
```

### Column {width=42%}

```{r}
#| title: "Test an ask: 45 h"
#| content: valuebox
#| color: "#1e3a28"
list(icon = "check-circle", value = f45$verdict)
```

```{r}
#| title: "How the 45 h ask is sourced"
src <- data.frame(
  Source = c("Under-run slack", "Contingency reserve", "Residual (change order)"),
  Hours  = c(f45$from_slack, f45$from_cont, f45$shortfall))
kable(src, align = "lr", row.names = FALSE)
```

Net findable = (planned + contingency) − EAC = (`r m$planned` + `r m$contingency`) − `r m$eac` = **`r m$findable` h**. The 45 h ask is absorbed with `r f45$contingency_left` h of reserve still held.

## Row {height=30%}

```{r}
#| title: "Effort ledger"
lg <- d$ledger
lg$slack <- pmax(0, lg$planned - lg$eac)
kable(lg[, c("deliverable","planned","actual","pct","eac","slack")],
      col.names = c("Deliverable","Planned","Actual","% done","EAC","Under-run slack"),
      row.names = FALSE, align = "lrrrrr")
```
