---
title: "Find Hours"
subtitle: "Budget reconciliation, effort only"
format:
  revealjs:
    theme: dark
    embed-resources: true
    slide-number: true
    smaller: true
    transition: fade
    fig-width: 9
    fig-height: 4.2
mermaid-format: svg
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(ggplot2); library(knitr)
source("../shiny/R/logic.R")
d   <- hours_load()
m   <- hours_metrics(d$ledger, d$contingency)
f45 <- find_hours(45, d$ledger, d$contingency)
f90 <- find_hours(90, d$ledger, d$contingency)
```

## The one-line idea {.scrollable}

When a new out-of-scope ask lands, **find the hours you already have** before asking for more.

$$\textbf{Net findable} = \big(\textstyle\sum\text{Planned} + \text{Contingency}\big) - \textstyle\sum\text{EAC}$$

- **Under-run slack**: planned effort you are forecast not to spend (planned − EAC, floored at 0)
- **Contingency**: the management reserve

Only the residual becomes a change order.

*Hours / effort only. Synthetic data. We say participant, not subject.*

## Where the numbers come from {.scrollable}

| Column | Source of record |
|---|---|
| Planned | SOW / budget grid in CTMS |
| Actual to date | Timesheet system |
| % complete, EAC | Your estimate + progress tracker |
| Change order | Contract: lead biostat + PM review, sponsor sign-off, SOW amended |

The tool never invents a number. You transcribe current figures in, so the reconciliation is attributable.

## What to do with a new ask {.scrollable}

```{mermaid}
%%{init: {'theme':'base','themeVariables':{'primaryColor':'#28344f','primaryTextColor':'#eef2fb','primaryBorderColor':'#7d9bd6','lineColor':'#aab8d6','tertiaryColor':'#20406a','tertiaryTextColor':'#eef2fb','tertiaryBorderColor':'#7d9bd6','secondaryColor':'#20406a','secondaryTextColor':'#eef2fb'}}}%%
flowchart LR
  ASK["New ask:<br/>estimate effort"] --> SLACK{"Under-run slack<br/>covers it?"}
  SLACK -->|Yes| REALLOC["Reallocate<br/>slack"]
  SLACK -->|Partly / No| CONT{"Contingency<br/>covers rest?"}
  CONT -->|Yes| DRAW["Draw the<br/>reserve"]
  CONT -->|Partly / No| DEFER{"Defer /<br/>re-sequence?"}
  DEFER -->|Yes| RESEQ["Free planned<br/>hours"]
  DEFER -->|No| CO["Change order:<br/>PM + sponsor"]
  REALLOC --> ABS["Absorbed"]
  DRAW --> ABS
  RESEQ --> ABS
```

## The live roll-up {.scrollable}

```{r kpis, results='asis'}
box <- function(label, value, col) sprintf(
  '<div style="flex:1;min-width:120px;background:#222630;border:1px solid #3a4157;border-radius:10px;padding:14px 16px;margin:4px"><div style="font-size:30px;font-weight:700;color:%s;font-family:monospace">%s</div><div style="font-size:12px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em">%s</div></div>', col, value, label)
cat('<div style="display:flex;flex-wrap:wrap;gap:2px">')
cat(box("Planned",     m$planned,     "#e9edf7"))
cat(box("Spent",       m$actual,      "#5b8cff"))
cat(box("EAC",         m$eac,         "#e0a64c"))
cat(box("Contingency", m$contingency, "#3fd0c9"))
cat(box("Net findable",m$findable,    "#43c47d"))
cat('</div>')
```

<p style="margin-top:18px;font-size:22px">
( `r m$planned` + `r m$contingency` ) − `r m$eac` = <b style="color:#43c47d">`r m$findable` h</b> you can redirect before a change order.
</p>

## Planned vs forecast {.scrollable}

```{r plot}
df <- d$ledger[, c("deliverable","planned","eac")]
long <- rbind(
  data.frame(deliverable = df$deliverable, hours = df$planned, kind = "Planned"),
  data.frame(deliverable = df$deliverable, hours = df$eac,     kind = "Forecast EAC"))
long$deliverable <- factor(long$deliverable, levels = rev(df$deliverable))
ggplot(long, aes(hours, deliverable, fill = kind)) +
  geom_col(width = 0.7, position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = c("Planned"="#5b8cff","Forecast EAC"="#e0a64c")) +
  labs(x = "Hours", y = NULL, fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top",
        plot.background  = element_rect(fill="#222630", colour=NA),
        panel.background = element_rect(fill="#222630", colour=NA),
        text = element_text(colour="#e9edf7"),
        axis.text = element_text(colour="#c5cee4"),
        panel.grid = element_line(colour="#3a4157"))
```

## Ask 45 h, absorbable {.scrollable}

```{r w45, results='asis'}
vbox <- function(fh){
  ok <- fh$absorbable
  bg <- if(ok)"#1e3a28" else "#3a1e24"; bd <- if(ok)"#43c47d" else "#e8657a"
  cat(sprintf('<div style="background:%s;border:1px solid %s;border-radius:12px;padding:20px 24px;max-width:720px;font-size:22px">', bg, bd))
  cat(sprintf('<div style="font-size:28px;font-weight:700;margin-bottom:12px">Ask %d h &rarr; %s</div>', fh$need, fh$verdict))
  cat(sprintf('<div style="font-family:monospace"><span style="color:#43c47d">%d h</span> under-run slack + <span style="color:#3fd0c9">%d h</span> contingency', fh$from_slack, fh$from_cont))
  cat(sprintf(' + <span style="color:%s">%d h</span> residual</div>', if(fh$shortfall>0)"#e8657a" else "#94a1c0", fh$shortfall))
  cat(sprintf('<div style="font-size:15px;color:#94a1c0;margin-top:10px">Contingency left: %d h</div></div>', fh$contingency_left))
}
vbox(f45)
```

Fully covered from what you already hold: no residual, `r f45$contingency_left` h of reserve still in hand.

## Ask 90 h, change order {.scrollable}

```{r w90, results='asis'}
vbox(f90)
```

Slack and the full reserve are exhausted. The **`r f90$shortfall` h** residual is genuine new scope: it goes to the change-order process. The tool drafts the hours-only note; pricing is handled downstream by the PM.

## Run it / build it {.scrollable}

```r
shiny::runApp("budget_reconciliation/shiny")
# install.packages(c("shiny","bslib","DT","ggplot2"))
```

- **Run**: [Run &amp; build the Shiny app](hours-build.html), edit an EAC, change the ask, watch net-findable recompute
- **Share**: [dashboard](hours-dashboard.html) · **Skim**: [one-pager](hours-onepager.pdf)
- **File**: [Excel workbook](../Hours_Budget_Reconciliation_Workbook.xlsx)
- **Build**: [build-it walkthrough](hours-build.html)

*Synthetic data. Hours / effort only. A change order is a named-human, contracted decision.*
