---
title: "Run and build the find-hours app"
subtitle: "The complete R Shiny app, every line"
format:
  html:
    theme: darkly
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    highlight-style: github-dark
    code-copy: true
    code-fold: false
execute:
  eval: false
  echo: true
---

Hours / effort only throughout: no dollars, rates, or headcount. The one equation:

```
NET FINDABLE = (Sum of Planned + Contingency) - Sum of EAC
```

## Run it

The interactive tool is a real R Shiny app. From the package root:

```r
install.packages(c("shiny", "bslib", "DT", "ggplot2"))   # once
shiny::runApp("budget_reconciliation/shiny")
```

Edit a task's EAC, change the ask, and the net-findable roll-up and the
absorbable-vs-change-order verdict recompute live. Download the change-order draft when
there is a shortfall, and the reconciled ledger to CSV for the [Excel workbook](../Hours_Budget_Reconciliation_Workbook.xlsx).

The whole app is two files, shown here in full, nothing elided:

- **shiny/R/logic.R** &mdash; the pure reconciliation math (no Shiny). The [visual guide](hours-guide.html), the [dashboard](hours-dashboard.html), and the unit tests call these same functions.
- **shiny/app.R** &mdash; the UI and the reactive server.

The one idea to hold onto: the **rules** live in `logic.R`, the **screen** lives in `app.R`. You can change what "net findable" means without touching the app, and change how it looks without touching the rules.

## The data it reads

One CSV is the whole dataset. Edit it to point the app at your study (keep the column names): [`project_hours_ledger.csv`](../templates/project_hours_ledger.csv). The rows named `CONTINGENCY` and `TOTAL` are handled specially by `hours_load()`: the contingency row sets the reserve, the total row is dropped, and everything else is a real deliverable.

## The complete logic core: shiny/R/logic.R

Every line of the logic. `hours_load()` reads the ledger CSV and splits out the
contingency reserve; `hours_metrics()` is the portfolio roll-up; `find_hours()` sources a
new ask from under-run slack first, then contingency, and reports the residual;
`change_order_text()` drafts the hours-only note. Nothing here depends on Shiny, so you can
test it in a plain console.

```{r logic, code=readLines("../shiny/R/logic.R"), eval=FALSE, echo=TRUE}
```

Read `find_hours()`: it sources the ask in a fixed order. `from_slack <- min(need, m$net_slack)`
takes what it can from under-run slack, `from_cont <- min(need - from_slack, contingency)`
takes the rest from the reserve, and `shortfall <- need - from_slack - from_cont` is
whatever is left. The ask is absorbable exactly when `shortfall <= 0`. Change that sourcing
order and you have changed what "find the hours you already have" means, everywhere, at once.

## The complete app: shiny/app.R

Every line of the UI and server. Read it top to bottom; the walk-through below points
at the parts that matter.

```{r app, code=readLines("../shiny/app.R"), eval=FALSE, echo=TRUE}
```

## How the reactivity works, part by part

Follow the data, not the layout. Here is the chain that makes a click flip the verdict.

**The editable state lives in one place.** `rv <- reactiveValues(ledger = withVar(seed$ledger))`
holds the one frame the user can change. The helper `withVar()` recomputes the `variance`
column (`planned - eac`) each time the ledger is rebuilt. Every output reads from `rv`, so
when `rv` changes, everything that depends on it recomputes. That is the entire trick.

**The rules are reactive expressions.** `metrics <- reactive(hours_metrics(rv$ledger, input$contingency))`
and `fh <- reactive(find_hours(input$need, rv$ledger, input$contingency))` re-run automatically
whenever the ledger **or** the ask input **or** the contingency changes. You never call the
roll-up by hand.

**The value boxes just read the reactive.** `output$v_planned`, `output$v_actual`,
`output$v_eac`, `output$v_cont`, and `output$v_find` each render one field of `metrics()`.
They know nothing about how it was computed.

**Editing a cell mutates the state.** The ledger is a `DT` with per-cell editing enabled on
the numeric columns. The observer on `input$ledger_cell_edit` coerces the new value with
`as.numeric()`, writes it into `rv$ledger` for that row and column, recomputes the `variance`
column, and pushes the frame back through a `dataTableProxy` with `replaceData()` so the user
does not lose their place. That write is what makes the verdict flip, because `fh()` depends
on `rv$ledger`.

**The quick-pick buttons set the ask.** Each of `input$q20`, `input$q45`, `input$q90`, and
`input$q150` calls `updateNumericInput(session, "need", value = 90)` (or 20, 45, 150), which retriggers `fh()`.
The `input$reset` observer restores `rv$ledger` from the seed.

**The verdict reads the reactive and branches.** `output$verdict` reads `fh()`, shows the
three-line sourcing breakdown (slack, contingency, shortfall), then renders a green
"Absorbable internally" alert when `f$absorbable` is true, or a red "Change order needed"
alert with the residual hours when it is not.

**The downloads round-trip to Excel and the eTMF.** `output$dl_ledger` writes `rv$ledger`
to CSV with the workbook's column names; `output$dl_co` writes `change_order_text(fh())`,
the hours-only change-order draft, to a Markdown file.

## Verify it without a browser

You can drive the reactive graph headless, which is how this app was checked before it
was ever opened in a browser:

```r
shiny::testServer("budget_reconciliation/shiny", {
  # a 90 h ask exhausts slack and reserve, leaving a change-order residual
  session$setInputs(need = 90, contingency = 40)
  fh()$shortfall            # 31

  # a 45 h ask is fully absorbable
  session$setInputs(need = 45)
  fh()$absorbable           # TRUE
})
```

## The unit tests: shiny is not involved

The reconciliation logic is covered by a zero-dependency test you run with `Rscript`. It is
the same file the guide and dashboard trust:

```{r tests, code=readLines("../tests/test-logic.R"), eval=FALSE, echo=TRUE}
```

Run it:

```r
Rscript budget_reconciliation/tests/test-logic.R   # prints: find-hours logic: all tests passed
```

## Make it yours

- Change the data: edit `project_hours_ledger.csv` in `templates/`.
- Change the rule: edit `hours_metrics()` or `find_hours()` in `shiny/R/logic.R`, for example to hold back part of the contingency as untouchable.
- Add a source step: insert it into the sourcing order in `find_hours()`, before the shortfall line.
- Add a column: add it to the CSV and to the `datatable()` column list in `app.R`.

The same skeleton, a logic core plus `reactiveValues` plus reactive rules plus outputs,
builds the [UAT tracker](../../uat_irt/quarto/uat-build.html) and the four dashboards too.
