---
title: "Run and build the Study Knowledge Graph"
subtitle: "The complete R Shiny graph, every line"
format:
  html:
    theme: darkly
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    highlight-style: github-dark
    code-copy: true
    code-fold: false
execute:
  eval: false
  echo: true
---

## Run it

The interactive tool is a real R Shiny app. From the package root:

```r
install.packages(c("shiny", "bslib", "visNetwork", "igraph"))   # once
shiny::runApp("knowledge_vault/shiny")
```

Filter by note type, click a note to read its detail and its neighbours, and read the
graph metrics (notes, links, clusters, the hub). Synthetic study, no PHI.

The whole app is two files, shown here in full, nothing elided:

- **shiny/R/logic.R** &mdash; the pure graph logic (no Shiny). The [visual guide](kg-guide.html), the [interactive graph](kg-graph.html), and the unit tests call these same functions.
- **shiny/app.R** &mdash; the UI and the reactive server that draws the graph.

The one idea to hold onto: the **graph math** lives in `logic.R`, the **screen** lives in `app.R`. You can change what "important" means without touching the app, and change how it looks without touching the math.

## The data it reads

Three CSVs are the whole vault. Edit them to point the app at your study (keep the column names): [`nodes.csv`](../shiny/data/nodes.csv), [`edges.csv`](../shiny/data/edges.csv), [`types.csv`](../shiny/data/types.csv).

```
nodes.csv:  id, title, type, status, owner, cluster, summary
edges.csv:  from, to
types.csv:  type, label, color
```

Each row of `nodes.csv` is one study artifact (protocol, SAP, ADaM specs, DMP, monitoring plan, milestones). Each row of `edges.csv` is one dependency between two notes. `types.csv` gives every note type a label and a colour. Nothing about the study is hard-coded.

## The complete logic core: shiny/R/logic.R

Every line of the graph logic. `kg_load()` reads the three CSVs; `kg_igraph()` builds
the undirected `igraph`; `kg_metrics()` attaches degree and betweenness to each note;
`kg_neighbors()` lists what a note links to; `kg_stats()` is the value-box row (notes,
links, clusters, the hub, the top broker); `kg_vis()` shapes nodes and edges for
`visNetwork`. Nothing here depends on Shiny, so you can test it in a plain console.

```{r logic, code=readLines("../shiny/R/logic.R"), eval=FALSE, echo=TRUE}
```

Read the metrics: `kg_metrics()` calls `igraph::degree()` (how many notes link to a
note, which drives node size) and `igraph::betweenness()` (how much a note sits
*between* clusters, which finds the broker). `kg_stats()` then sorts by each to pick
the hub and the broker. Those numbers are computed from the edges, never asserted.

## The complete app: shiny/app.R

Every line of the UI and server. Read it top to bottom; the walk-through below points
at the parts that matter.

```{r app, code=readLines("../shiny/app.R"), eval=FALSE, echo=TRUE}
```

## How the reactivity works, part by part

Follow the data, not the layout. Here is the chain that makes a checkbox redraw the
graph and a click fill the detail panel.

**The graph is loaded once, up front.** `kg <- kg_load("data")` and `V <- kg_vis(kg)`
run when the app starts, so the shaped nodes and edges (`V$nodes`, `V$edges`) are ready
before any input. The value-box numbers come from `st <- kg_stats(kg)` in the server.

**The type checkboxes drive one reactive.** The `sub` reactive keeps only the
nodes whose `group` is in `input$types`, then keeps only the edges whose both ends
survive. Re-checking a type re-runs `sub()`, which is the only thing `renderVisNetwork`
reads, so the subgraph redraws. You never redraw by hand.

**Drawing the graph reads that reactive.** `output$net` renders the `visNetwork` from
`sub()` and attaches a `selectNode` event handler. That JavaScript callback fires
`Shiny.setInputValue("kg_sel", nodes.nodes[0])` whenever you click a node, which is how
a click in the browser becomes an R input.

**Selecting a note fills the detail panel.** `selected` is a `reactiveVal` that starts
at `"hub"`. `observeEvent(input$kg_sel, selected(input$kg_sel))` updates it on a click,
and the focus dropdown updates it too. `output$detail` reads `selected()`, looks up that
note, and calls `kg_neighbors(kg, id)` to list what it links to.

**Focus just recenters.** The `input$focus` observer uses a `visNetworkProxy` to
call `visFocus()` and `visSelectNodes()` on the chosen note, without redrawing the whole
graph.

**Everything else just reads the loaded data.** The value boxes read `st`, the graph
reads `sub()`, and the detail panel reads `selected()` plus `kg_neighbors()`. Change the
CSVs and every one of these updates on the next run.

## Verify it without a browser

The graph logic has no Shiny in it, so you can drive it in a plain console. This is how
the logic was checked before it was ever opened in a browser:

```r
source("knowledge_vault/shiny/R/logic.R")
kg <- kg_load("knowledge_vault/shiny/data")
st <- kg_stats(kg)
st$n_nodes           # number of notes
st$n_edges           # number of links
st$hub_title         # the most-connected note
st$broker_title      # the note that most brokers between clusters
nrow(kg_neighbors(kg, "sap"))   # how many notes the SAP links to
```

## The unit tests: shiny is not involved

The graph logic is covered by a zero-dependency test you run with `Rscript`. It is the
same file the guide and the interactive graph trust:

```{r tests, code=readLines("../tests/test-logic.R"), eval=FALSE, echo=TRUE}
```

Run it:

```r
Rscript knowledge_vault/tests/test-logic.R   # prints: Knowledge-graph logic: all tests passed
```

## Make it yours

- **Add a note:** a row in [`nodes.csv`](../shiny/data/nodes.csv) plus its links in [`edges.csv`](../shiny/data/edges.csv). It appears in the graph, sized by how many notes link to it.
- **Add a category:** a row in [`types.csv`](../shiny/data/types.csv) (type, label, colour), and it becomes a checkbox and a legend entry.
- **Change what "important" means:** size nodes by `betweenness` instead of `degree` in `kg_vis()`. Nothing else changes, because everything reads that one function.
- **Add a panel:** another `card()` in the UI and a matching `render*` in the server.

The same skeleton, a logic core plus data as CSV plus one render path, builds the
[lifecycle monitor](../../lifecycle_dashboard/quarto/lifecycle-build.html) and the
[UAT tracker](../../uat_irt/quarto/uat-build.html) too.
