---
title: "Trial-Termination Early-Warning - Snapshot"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    embed-resources: true
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr)
source("../shiny/R/logic.R")
s <- risk_load("../shiny/data")
o <- risk_overall(s)
```

::: {.callout-note appearance="simple"}
**Read-only shareable snapshot**, computed from the signal feed [`signals.csv`](../shiny/data/signals.csv). The interactive tool is the [Run &amp; build the Shiny app](risk-build.html) page, and the [visual guide](risk-guide.html) explains the roll-up. The AI flags; the SRC / DSMB / medical monitor / PM / account lead decide. Synthetic, no PHI, no participant records.
:::

```{r kpis, results='asis'}
box <- function(label, score, rag) {
  col <- c(green="#43c47d", amber="#e0a64c", red="#e8657a")[rag]
  sprintf('<div style="flex:1 1 150px;background:#222630;border:1px solid #3a4157;border-radius:10px;padding:14px 16px"><div style="font-size:26px;font-weight:700;color:%s;font-family:monospace;line-height:1">%d · %s</div><div style="font-size:11px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em;margin-top:6px">%s</div></div>',
    col, score, toupper(rag), label)
}
cat('<div style="display:flex;flex-wrap:wrap;gap:10px;margin:6px 0 18px">')
cat(box("Overall termination risk", o$score, o$rag))
for (t in TIERS) { ts <- o$tiers[[t]]; cat(box(paste(t, "tier"), ts$score, ts$rag)) }
cat('</div>')
```

:::: {.grid}

::: {.g-col-12 .g-col-md-7}
#### All signals
```{r table}
d <- s[, c("tier","id","signal","category","engine","level","weight")]
d$adv <- ifelse(as.logical(s$advanced), "ADV", "")
kable(d, col.names=c("Tier","ID","Signal","Category","Engine","Level","Wt","Adv"), row.names=FALSE)
```
:::

::: {.g-col-12 .g-col-md-5}
#### Early-warning feed (firing, ranked)
```{r feed, results='asis'}
f <- risk_feed(s)
for (j in seq_len(nrow(f))) {
  r <- f[j, ]; col <- if (r$level == "red") "#e8657a" else "#e0a64c"
  adv <- if (as.logical(r$advanced)) ' <span style="font-size:10px;background:#3a4157;border-radius:4px;padding:1px 5px">ADV</span>' else ""
  cat(sprintf('<div style="display:flex;gap:8px;padding:6px 0;border-bottom:1px solid #2b3350"><div style="flex:0 0 8px;height:8px;border-radius:50%%;background:%s;margin-top:6px"></div><div><div style="font-size:13px;color:#e9edf7"><b>%s</b> · <span style="font-family:ui-monospace,monospace;color:#7fb0ff">%s</span> %s%s</div><div style="font-size:12px;color:#94a1c0">%s</div></div></div>',
    col, r$tier, r$id, r$signal, adv, r$detail))
}
```
:::

::::

*The advanced-engine signals (marked ADV) use published methods: Optimal Transport / Wasserstein drift and Active Inference / free-energy. Reported safety numbers come from the validated safety database, never a model.*
