---
title: "Trial-Termination Early-Warning"
subtitle: "A visual, runnable guide to the 3-tier risk console"
format:
  html:
    theme:
      - darkly
      - ../../_quarto-mc.scss
    include-before-body: ../../_mc-topbar.html
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    code-tools: true
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
---

```{r setup}
library(knitr)
source("../shiny/R/logic.R")
s <- risk_load("../shiny/data")
o <- risk_overall(s)
```

```{=html}
<style>
svg[id^="mermaid"]{ width:100% !important; max-width:940px; height:auto; display:block; margin:8px auto; }
svg[id^="mermaid"] .nodeLabel, svg[id^="mermaid"] .edgeLabel { font-size:16px !important; }
</style>
```

::: {.callout-note appearance="simple"}
**Run / share / read / build - pick your mode.** This page *shows* the console (diagrams + a live roll-up), the [**Run &amp; build the Shiny app**](risk-build.html) lets you *run* it and set a what-if, the [dashboard](risk-dashboard.html) is a shareable snapshot, and the [run-and-build page](risk-build.html) shows how it is made. Synthetic, no PHI, no participant records.
:::

::: {.callout-important appearance="simple"}
**The rule.** The AI **flags, aggregates, and drafts** the alert. The SRC / DSMB / medical monitor make every DLT and safety call; the PM owns the study; the account lead owns the client signal. Reported safety numbers come from the validated safety database, never from a model.
:::

## Three tiers, one roll-up

```{mermaid}
%%{init: {'theme':'base','themeVariables':{'primaryColor':'#28344f','primaryTextColor':'#eef2fb','primaryBorderColor':'#7d9bd6','lineColor':'#aab8d6','clusterBkg':'#161d33','clusterBorder':'#3a476a','tertiaryColor':'#20406a','tertiaryTextColor':'#eef2fb','tertiaryBorderColor':'#7d9bd6','secondaryColor':'#20406a','secondaryTextColor':'#eef2fb'}}}%%
flowchart LR
  SRC["Data sources<br/>EDC · safety DB · labs<br/>IXRS · CTMS · finance"] --> RT{"Route by<br/>sensitivity"}
  RT -->|"safety / PK /<br/>unblinded"| LOC["On-device model<br/>zero egress"]
  RT -->|"non-sensitive<br/>ops"| CLD["Cloud Claude<br/>API + BAA"]
  LOC --> ADV["Advanced engine<br/>OT + Active Inference"]
  CLD --> ADV
  ADV --> SC["Severity-weighted<br/>roll-up to tier RAG"]
  SC --> ESC["Alert +<br/>escalation"]
```

The scoring is transparent - each signal has a level (green/amber/red) and a
weight; a tier is the weighted blend, and the overall index weights the tiers
(participant safety heaviest).

## The live console - computed from the signal feed {#live}

Everything below is computed from one file, the signal feed: [`signals.csv`](../shiny/data/signals.csv) (in the [`shiny/data/`](../shiny/data/) folder). Edit that CSV (or run the app) and the roll-up, the tiers, and the feed all move.

```{r kpis, results='asis'}
box <- function(label, score, rag) {
  col <- c(green="#43c47d", amber="#e0a64c", red="#e8657a")[rag]
  sprintf('<div style="flex:1 1 150px;background:#222630;border:1px solid #3a4157;border-radius:10px;padding:14px 16px"><div style="font-size:26px;font-weight:700;color:%s;font-family:monospace;line-height:1">%d · %s</div><div style="font-size:11px;color:#94a1c0;text-transform:uppercase;letter-spacing:.05em;margin-top:6px">%s</div></div>',
    col, score, toupper(rag), label)
}
cat('<div style="display:flex;flex-wrap:wrap;gap:10px;margin:6px 0 18px">')
cat(box("Overall termination risk", o$score, o$rag))
for (t in TIERS) { ts <- o$tiers[[t]]; cat(box(paste(t, "tier"), ts$score, ts$rag)) }
cat('</div>')
```

::: {.panel-tabset}

### Early-warning feed

```{r feed, results='asis'}
f <- risk_feed(s)
for (j in seq_len(nrow(f))) {
  r <- f[j, ]; col <- if (r$level == "red") "#e8657a" else "#e0a64c"
  adv <- if (as.logical(r$advanced)) ' <span style="font-size:10px;background:#3a4157;border-radius:4px;padding:1px 5px">ADV</span>' else ""
  cat(sprintf('<div style="display:flex;gap:8px;padding:6px 0;border-bottom:1px solid #2b3350"><div style="flex:0 0 8px;height:8px;border-radius:50%%;background:%s;margin-top:6px"></div><div><div style="font-size:13px;color:#e9edf7"><b>%s</b> · <span style="font-family:ui-monospace,monospace;color:#7fb0ff">%s</span> %s%s</div><div style="font-size:12px;color:#94a1c0">%s</div></div></div>',
    col, r$tier, r$id, r$signal, adv, r$detail))
}
```

### All signals

```{r table}
d <- s[, c("tier","id","signal","category","engine","level","weight")]
kable(d, col.names=c("Tier","ID","Signal","Category","Engine","Level","Weight"), row.names=FALSE)
```

:::

## Run it - and change it {#run}

```r
shiny::runApp("risk_monitor/shiny")
# needs: install.packages(c("shiny","bslib","DT"))
```

Select a signal, set its level, and watch the tier and overall risk re-colour. The
[run-and-build page](risk-build.html) shows how the console is built from `R/logic.R`.

::: {.callout-tip appearance="simple"}
**Going live:** the feed file [`signals.csv`](../shiny/data/signals.csv) is a representative subset of the 65-signal catalog. Replace it with your KRI / stopping-rule feed (same columns) and the roll-up is yours. Participant-tier signals are de-identified states, never records.
:::

## The other formats

| Mode | Artifact |
|---|---|
| Run | [Run &amp; build the Shiny app](risk-build.html) |
| Share (no R) | [Dashboard](risk-dashboard.html) |
| Build | [Run-and-build walkthrough](risk-build.html) |

*Synthetic, no PHI. The AI flags; the SRC / DSMB / PM / account lead decide.*
