---
title: "Run and build the Trial-Termination Early-Warning console"
subtitle: "The complete R Shiny console, every line"
format:
  html:
    theme: darkly
    toc: true
    toc-location: left
    toc-title: On this page
    embed-resources: true
    highlight-style: github-dark
    code-copy: true
    code-fold: false
execute:
  eval: false
  echo: true
---

## Run it

The interactive tool is a real R Shiny app you own. From the package root:

```r
install.packages(c("shiny", "bslib", "DT"))   # once
shiny::runApp("risk_monitor/shiny")
```

Select a signal, set its level (green / amber / red), and watch the tier RAG and the
overall termination-risk index re-colour. Use "Reset to feed" to return to the seed
state. The AI flags and drafts; the SRC / DSMB / medical monitor / PM / account lead
decide. Synthetic, no PHI, no participant records.

The whole console is two files, shown here in full, nothing elided:

- **shiny/R/logic.R** &mdash; the pure roll-up scoring (no Shiny). The [visual guide](risk-guide.html), the [dashboard](risk-dashboard.html), and the unit tests call these same functions.
- **shiny/app.R** &mdash; the UI and the reactive server.

The one idea to hold onto: the **rules** live in logic.R, the **screen** lives in app.R. You can change what a tier RAG means without touching the app, and change how it looks without touching the rules.

## The data it reads

One CSV is the whole dataset. Edit it to point the console at your study (keep the column names): [`signals.csv`](../shiny/data/signals.csv). It carries `tier, id, signal, category, engine, level, weight, advanced, detail`, a representative subset of the 65-signal catalog. The `advanced = TRUE` rows are the continuous signals from the published-method engine (Optimal Transport / Wasserstein drift, Active Inference / free-energy).

## The complete logic core: shiny/R/logic.R

Every line of the roll-up. `risk_load()` reads the CSV and levels the RAG factor; `tier_score()` is the severity-weighted blend of one tier; `risk_overall()` weights the tiers (participant safety heaviest); `risk_feed()` is the firing-only, red-first list; `risk_counts()` is the tally. Nothing here depends on Shiny, so you can test it in a plain console.

```{r logic, code=readLines("../shiny/R/logic.R"), eval=FALSE, echo=TRUE}
```

Read the roll-up: each signal has a level mapped to a severity in `SEVERITY` (green 0, amber 50, red 100) and a per-signal `weight`. A tier score is the weight-weighted mean of its signals' severities, and `risk_overall()` blends the three tier scores with `TIER_WEIGHT` (Participant 0.50, Study 0.30, Client 0.20). `rag_of()` turns any score into a colour at the thresholds 25 and 55. Change a weight or a threshold on that one line and every surface, the guide, the dashboard, and the app, updates at once.

## The complete app: shiny/app.R

Every line of the UI and server. Read it top to bottom; the walk-through below points at the parts that matter.

```{r app, code=readLines("../shiny/app.R"), eval=FALSE, echo=TRUE}
```

## How the reactivity works, part by part

Follow the data, not the layout. Here is the chain that makes a click re-colour the risk index.

**The editable state lives in one place.** `rv <- reactiveValues(signals = seed)` holds the one frame the user can change. Every output reads from `rv$signals`, so when it changes, everything that depends on it recomputes. That is the entire trick.

**The what-if writes into that state.** The observer on `input$level` finds the selected signal by `input$sig`, and if the chosen level differs, it writes a new `factor` into `rv$signals$level` for that row. Selecting a row in the table calls `updateSelectInput` so the sidebar `input$sig` follows your click, and switching `input$sig` calls `updateRadioButtons` so the level control shows that signal's current level.

**The rule is a reactive expression.** `ov <- reactive(risk_overall(rv$signals))` re-runs automatically whenever `rv$signals` changes. You never call `risk_overall()` by hand.

**The value boxes just read the reactive.** The helper `vb()` reads `ov()`, pulls either the overall score or a named tier from `o$tiers`, colours the span by its RAG, and each of `output$v_overall`, `output$v_part`, `output$v_study`, and `output$v_client` renders it. When `rv$signals` changes, `ov()` recomputes and all four boxes re-colour.

**The table and the feed read the state directly.** `output$signals` renders `rv$signals` (filtered by `input$tier`) as a single-select `DT`; `output$feed` calls `risk_feed(rv$signals)` and lays out the firing signals red-first, badging the `advanced` ones "ADV". Nothing recomputes by hand; changing `rv$signals` drives it all.

**Reset restores the seed.** The observer on `input$reset` assigns `rv$signals <- seed` and re-points the level control at the seed level of the current signal.

## Verify it without a browser

You can drive the reactive graph headless, which is how this console was checked before it was ever opened in a browser:

```r
shiny::testServer("risk_monitor/shiny", {
  # escalate a Study-tier signal and watch that tier rise
  session$setInputs(sig = "ESC-33", level = "red")
  ov()$tiers$Study$score      # higher than the seed value
  ov()$score                  # the overall index moves with it
})
```

## The unit tests: shiny is not involved

The roll-up is covered by a zero-dependency test you run with `Rscript`. It is the same file the guide and dashboard trust:

```{r tests, code=readLines("../tests/test-logic.R"), eval=FALSE, echo=TRUE}
```

Run it:

```r
Rscript risk_monitor/tests/test-logic.R      # prints: Risk logic: all tests passed
```

## Make it yours

- **Change the rule:** edit `rag_of()` or the `TIER_WEIGHT` weights in `shiny/R/logic.R`. Nothing else changes, because everything calls those functions.
- **Change the data:** swap `data/signals.csv` for your live KRI / stopping-rule feed (same columns).
- **Add a tier or a signal:** add rows to the CSV and, for a new tier, add a `TIER_WEIGHT` entry and a value box in `app.R`.

The same skeleton, a logic core plus `reactiveValues` plus a reactive rule plus outputs, builds the [lifecycle monitor](../../lifecycle_dashboard/quarto/lifecycle-build.html), the [UAT tracker](../../uat_irt/quarto/uat-build.html), and the [find-hours app](../../budget_reconciliation/quarto/hours-build.html) too.
